Ownership types in theory and practice (in Rust)

### 2023-01-23 slides/slides.tex
Added third will it compile example

\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{listings-rust}
\lstset{
  frame=tb,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle=\footnotesize\ttfamily,
  % basicstyle={\small\ttfamily},
  numbers=left,
  numbersep=5pt,
  % xleftmargin=\parindent,
  numberstyle=\color{gray},
  keywordstyle=\bfseries\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=false,
  tabsize=2
}
\usepackage{tikz}
\usetikzlibrary{fit,positioning}
\definecolor{tumblue}{RGB}{0,101,189}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{cgreen}{RGB}{60,201,65}
\setbeamerfont{footnote}{size=\tiny}
\renewcommand\footnoterule{}
\usecolortheme{orchid}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{sidebar right}{% also implies no navbar
  \llap{\tikz\fill[tumblue,scale=.1] (0,0) ++(-5cm,-5cm) ++(-10cm,0)
      +(0cm,0cm) -- +(4cm,0cm) -- +(4cm,-4cm) -- +(5cm,-4cm) -- +(5cm,0cm) --
      +(10cm,0cm) -- +(10cm,-5cm) -- +(9cm,-5cm) -- +(9cm,-1cm) --
      +(8cm,-1cm) -- +(8cm,-5cm) -- +(7cm,-5cm) -- +(7cm,-1cm) --
      +(6cm,-1cm) -- +(6cm,-5cm) -- +(3cm,-5cm) -- +(3cm,-1cm) --
      +(2cm,-1cm) -- +(2cm,-5cm) -- +(1cm,-5cm) -- +(1cm,-1cm) --
      +(0cm,-1cm) -- cycle;}
  % Slide number in triangle.
  \vfill\llap{\tikz\path[fill=tumblue,inner sep=3pt] (0,0) -- ++(-1cm,0) -- +(1cm,1cm) -- cycle node[anchor=south east] {\usebeamerfont{footline}\color{white}\bfseries\insertframenumber};}%
  % Variant without triangle
  % \vfill\llap{\usebeamerfont{footline}\usebeamercolor[fg]{footline}\insertframenumber\hskip3pt}\vskip3pt%
}

@@

\begin{frame}[fragile]{Borrowing: examples}
\begin{itemize}
  \begin{lstlisting}[language=Rust]
  let mut s = String::from("hello");
  let b = &mut a;
  Vec::push(&mut v, 12);
  println!("{}", *b);
  \end{lstlisting}
  \pause

  \begin{lstlisting}[language=Rust]
  let mut v = vec![10, 11];
  Vec::push(&mut v, 12); // lifetime 'b starts and ends
  let vptr = &mut v[1];  // lifetime 'a starts
  println!("{}", *vptr); // lifetime 'a ends
  \end{lstlisting}

  \item Will it compile? \pause No.
  % TODO: lifetime explanation
  \item Why not? \pause Borrowed value 'b' does not live as long as the reference 'a'.
  \item Refactored version that compiles:\pause

>>>>

\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{listings-rust}
\lstset{
  frame=tb,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle=\footnotesize\ttfamily,
  numbers=left,
  numbersep=5pt,
  numberstyle=\color{gray},
  keywordstyle=\bfseries\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=false,
  tabsize=2
}
\usepackage{tikz}
\usetikzlibrary{fit,positioning}
\definecolor{tumblue}{RGB}{0,101,189}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{cgreen}{RGB}{60,201,65}
\setbeamerfont{footnote}{size=\tiny}
\renewcommand\footnoterule{}
\usecolortheme{orchid}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{sidebar right}{% also implies no navbar
  \llap{\tikz\fill[tumblue,scale=.1] (0,0) ++(-5cm,-5cm) ++(-10cm,0)
      +(0cm,0cm) -- +(4cm,0cm) -- +(4cm,-4cm) -- +(5cm,-4cm) -- +(5cm,0cm) --
      +(10cm,0cm) -- +(10cm,-5cm) -- +(9cm,-5cm) -- +(9cm,-1cm) --
      +(8cm,-1cm) -- +(8cm,-5cm) -- +(7cm,-5cm) -- +(7cm,-1cm) --
      +(6cm,-1cm) -- +(6cm,-5cm) -- +(3cm,-5cm) -- +(3cm,-1cm) --
      +(2cm,-1cm) -- +(2cm,-5cm) -- +(1cm,-5cm) -- +(1cm,-1cm) --
      +(0cm,-1cm) -- cycle;}
  % Slide number in triangle.
  \vfill\llap{\tikz\path[fill=tumblue,inner sep=3pt] (0,0) -- ++(-1cm,0) -- +(1cm,1cm) -- cycle node[anchor=south east] {\usebeamerfont{footline}\color{white}\bfseries\insertframenumber};}%
  % Variant without triangle
  % \vfill\llap{\usebeamerfont{footline}\usebeamercolor[fg]{footline}\insertframenumber\hskip3pt}\vskip3pt%
}

@@

\begin{frame}[fragile]{Borrowing: examples}
\begin{itemize}
  \begin{lstlisting}[language=Rust]
  let mut s = String::from("hello");
  let r = &s;
  s.push_str(" world");
  println!("{}", *r);
  \end{lstlisting}

  \item Will it compile? \pause No.
  \item Why not? \pause Lifetimes of mutable and immutable references overlap (cannot borrow 's' as mutable because it is also borrowed as immutable)
  \item Refactored version that compiles:\pause

  \begin{lstlisting}[language=Rust]
  let mut s = String::from("hello");
  s.push_str(" world"); // lifetime 'a starts and ends
  let r = &s; // lifetime 'b starts
  println!("{}", *r); // lifetime 'b ends
  \end{lstlisting}
\end{itemize}
\end{frame}
